### 2023-02-11 doc/manuscript/translocation.qmd
Add additional paragraphs to introduction

-   Evolutionary rescue be difficult to detect, and this complicates our understanding of the potential importance of evolutionary rescue in allowing the persistence and recovery of host populations following disease emergence (Eskew et al. 2021, Lilley et al. 2020). <!--# Maybe move this to the Discussion? -->

@@

P3: For resistant populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environments.

-   Translocation of individuals from putatively resistant <!--# Can we use "reduced susceptibilty" strategically to avoid potential pitfalls associated with resistance, tolerance, and our lack of understanding of what actual mechanism underlyies lower Bd loads and higher frog survival?  --> populations to sites that were formerly occupied but from which populations were extirpated by disease provides a way to disentangle the contributions of natal environment and host genetics on host persistence and recovery.

-   Something here about previous work with translocations and disease.

-   Simply put, resistant hosts that demonstrate disease resistance, population growth, and long-term viability in non-natal environments provide a means to recover disease-impacted species at a landscape-scale.

-   <!--# Zooming out here to make a broader point --> Few, if any, studies of disease-impacted species have provided comprehensive evidence that (1) evolutionary rescue can increase host resistance, (2) increased host resistance contributes to higher individual fitness (e.g., survival) in the presence of disease and allows population growth, (3) higher individual fitness and positive population growth applies beyond the natal environment, (4) individuals from higher-fitness populations can be used in translocations to reestablish extirpated populations, and (5) many reestablished populations have a high probability of persisting over the long term in a landscape of disease. <!--# I don't know how many of these boxes we can check, but they seem like important points to help us lay the foundation for our study (which checks more of these boxes than do most studies). I actually wonder if taking the broader approach outlined in this bullet would make for a more compelling argument than one focused more narrowly on issue #3, as most of the bullets in this paragraph currently do. That issue, although important, may distract from the broader perspective that I tried to highlight. -->

P4: Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery. Cite Knapp et al. (2016) to emphasize that frogs in recovering populations are more resistant to Bd than those in Bd-naive populations.

P5: Here, we combine 24 translocations of resistant MYL populations, X longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate that resistant populations of an endangered amphibian species can recover and persist across a landscape of disease.

We then use genetic evidence to demonstrate that observed disease resistance is likely a result of selection.

As such, we conclude that natural selection combined with translocations promotes the recovery of a species on the brink of disease-induced extinction. These results may have broad implications for the many species worldwide that are threatened with extinction by the emergence of novel diseases.

>>>>

-   Evolutionary rescue can be difficult to detect, and this complicates our understanding of the potential importance of evolutionary rescue in allowing the persistence and recovery of host populations following disease emergence (Eskew et al. 2021, Lilley et al. 2020). <!--# Maybe move this to the Discussion? -->

@@

<!--# P3: In this paragraph, I took a broader approach than what was originally outlined (which focused on fitness benefits extending beyond natal environment). I'm happy to revisit this approach as necessary. Also, can we use "reduced susceptibilty" strategically (especially in Results) to avoid potential pitfalls associated with resistance, tolerance, and our lack of understanding of what actual mechanism underlies lower Bd loads and higher frog survival?  -->

<!--# The following point was made in the original outline but is not included in P3, P4, or P5. Could be useful to include when introducing translocations in Results: Translocation of individuals from putatively resistant populations to sites that were formerly occupied but from which populations were extirpated by disease provides a way to disentangle the contributions of natal environment and host genetics on host persistence and recovery. -->

Given the dramatic impact of disease on global wildlife populations, reintroductions will often be necessary to reestablish extirpated populations at a landscape scale. However, there are several conditions that will often determine the feasibility of this much-needed recovery approach. First, populations that survived following disease arrival have evolved increased resistance. Second, increased resistance contributes to higher individual fitness in the presence of disease and allows positive population growth. Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment, allowing for the establishment of populations reintroduced into novel environments. And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease. To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021]. Furthermore, a lack of relevant information about disease or a lack of suitable options to overcome the threat of disease has resulted in the failure of numerous reintroduction efforts [@scheele2021; @fellers2007]. With the increase in emerging infectious diseases in recent decades [@daszak2000], these failures highlight the critical need to develop effective tools to recover disease-impacted biodiversity worldwide. <!--# Need to better tie these two sentences for the conditins listed above. -->

<!--# P4: Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery. -->

The mountain yellow-legged (MYL) frog, comprised of the sister species *Rana muscosa* and *Rana sierrae* [@vredenburg2007], is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus (*Batrachochytrium dendrobatidis*; [@scheele2019]). Once the most common amphibian in the high elevation portion of the Sierra Nevada mountains (California, USA; [@grinnell1924]), this frog has disappeared from more than 90% of its historical range during the past century [@vredenburg2007]. Due to the severity of its decline and the increasing probability of extinction, both species are now listed as "endangered" under the U.S. Endangered Species Act. Although this decline was initiated by the introduction of non-native trout into the once-fishless portion of the Sierra Nevada [@bradford1989; @knapp2000], the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread [@vredenburg2019] caused additional widespread population extirpations [@vredenburg2010; @rachowicz2006]. These Bd-caused declines are fundamentally different from the fish-caused declines because removal of fish populations is feasible and results in the rapid recovery of frogs. In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts [@walker2007], and therefore represents a permanent alteration of invaded ecosystems that amphibians will need to overcome to reestablish persistent populations.

<!--# P5: Here, we combine 24 translocations of resistant MYL populations, X longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate that resistant populations of an endangered amphibian species can recover and persist across a landscape of disease. We then use genetic evidence to demonstrate that observed disease resistance is likely a result of selection. As such, we conclude that natural selection combined with translocations promotes the recovery of a species on the brink of disease-induced extinction.  -->

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some frog populations have persisted following epizootics [@briggs2010] and are now recovering [@knapp2016]. Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant genotypes [@knapp2016]. Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018]. Here, we report the results of a greatly expanded MYL frog reintroduction effort and an associated analysis to evaluate genomic signatures of selection following MYL frog exposure to Bd. Collectively, the results show that MYL frogs generally meet the four conditions for reintroduction feasibility outlined above, and provide a novel example of how the reintroduction of resistant frogs can allow the large-scale recovery of this disease-impacted amphibian. These results may have broad implications for the many species worldwide that are threatened with extinction by emerging diseases.
